---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
---


## Loading and preprocessing the data

```{r,message=FALSE}
library(lubridate)
library(dplyr)
library(ggplot2)
library(lattice)
Sys.setlocale("LC_TIME", "en_US")
actWithNA<-read.csv('activity.csv')
actWithNA<-transform(actWithNA,date=ymd(date),steps=as.numeric(steps))
act<-na.omit(actWithNA)
```

## What is mean total number of steps taken per day?
```{r}
actSum<-act %>%
        group_by(date) %>%
        summarise(sumStep=sum(steps))
g<-ggplot(actSum,aes(x=sumStep))+
        geom_histogram(fill=colors()[55])
suppressMessages(print(g))
meanValue<-mean(actSum$sumStep)
medianValue<-median(actSum$sumStep)
meanValue
medianValue
```
The mean and median of the total number of steps taken per day is `r as.character(round(meanValue,2))` and `r as.character(medianValue)` respectively. 

## What is the average daily activity pattern?


```{r}
actDay<-act %>%
        group_by(interval) %>%
        summarise(averageStep=mean(steps))
plot(actDay,type='l')
whichInterval<-actDay$interval[which.max(actDay$averageStep)]
```
5-minute interval "`r whichInterval`"  contains the maximum number of steps on average across all the days in the dataset. 

## Imputing missing values

```{r}
 
naRow<-nrow(actWithNA)-sum(complete.cases(actWithNA))
```
`r naRow` rows with NA in the dataset
Imputing missing value by the mean value of the same interval as follow:
```{r}
imputedact<-merge(actWithNA,actDay,by='interval')
naId<-is.na(imputedact$steps)
imputedact$steps[naId]<-imputedact$averageStep[naId]
imputedactSum<-imputedact %>%
        group_by(date) %>%
        summarise(sumStep=sum(steps))
imputedg<-ggplot(imputedactSum,aes(x=sumStep))+
        geom_histogram(fill=colors()[55])
suppressMessages(print(imputedg))
imputedmeanValue<-mean(imputedactSum$sumStep)
imputedmedianValue<-median(imputedactSum$sumStep)
imputedmeanValue
imputedmedianValue
```
 
After imputing missing value by the mean value of the same interval. The mean and median of the total number of steps taken per day is `r as.character(round(imputedmeanValue,2))` and `r as.character(round(imputedmedianValue,2))` respectively. They seem very close to the estimates calculating from the dataset without imputation. These two figures shows the imputing missing data doesn't change the distribution.

## Are there differences in activity patterns between weekdays and weekends?
```{r}
weekDay<-weekdays(imputedact$date,abbreviate=T)
imputedact$isWeekend<-ifelse(weekDay=='Sat'|weekDay=='Sun','weekend','weekday')
imputedact$isWeekend<-as.factor(imputedact$isWeekend)
imputedactDay<-imputedact %>%
        group_by(interval,isWeekend) %>%
        summarise(averageStep=mean(steps))
xyplot(averageStep~interval|isWeekend,data=imputedactDay,type='l',
       layout = c(1, 2),
       ylab = "Number of steps")
```

